Fusion Moves for Correlation Clustering

### 2014-11-12 paper/egpaper_for_review.tex
reloacte section polyhedral representation

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0^{\breve{y}}  & =  \{ ij \in E \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0^{\breve{y}}$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0^{\breve{y}} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solution.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.

@@

\begin{theorem}[Equivalence]
The back projection of the optimal segmentation $\bar{y}'$ of the contracted graph $G_y=(V_y,E_y, \bar{w})$
is an optimal solution of problem~\ref{eq:fusion_move_a}.
\begin{proof}
Let $y'$ be the back propagation of $\bar{y}'$, which is by definition a feasible for~\ref{eq:fusion_move_a}.
If $y'$ would not be an optimal solution than there must exists $y''$ with 
$\sum_{e\in E} w_ey_e'' > \sum_{e\in E} w_ey_e''$.
Since $y_e'$ and $y_e''$ are $0$ for all $e \in \mathcal{E}_0^{\breve{y}}$ we would have
\begin{align*}
  \sum_{\bar{e}\in E_y} \bar{w}_{\bar{e}}\bar{y}'_{\bar{e}}
 &= \sum_{e\in E \setminus \mathcal{E}_0^{\breve{y}}} w_ey_e'          
  = \sum_{e\in E} w_ey_e'' \\
 &> \sum_{e\in E} w_ey_e''
  = \sum_{e\in E \setminus \mathcal{E}_0^{\breve{y}}} w_ey_e''
  =\sum_{\bar{e}\in E_y} \bar{w}_{\bar{e}}\bar{y}''_{\bar{e}}
\end{align*}
where $\bar{y}''$ is the projection from $y''$ on $G_y$.
This contradicts that $y'$ is a optimal segmentation of $G_y$.
\end{proof}
%
\end{theorem}

@@

%-------------------------------------------------------------------------
%\subsection{Fusion Move Solver}

@@

As discussed in \cite{Lempitsky-2010}, proposals
should have two properties: \emph{quality} 
and \emph{diversity}.
The quality / energy of the individual proposals
ensures low energy solutions since the
result of the fusion cannot be higher then
the energy of the current best solution.
Diversity between the individual proposals
reduces the chances to get stuck in local minima.
For correlation clustering fusion we add a third
property: \emph{size}.
The size of the contraction graph directly
depends on the number of connected components
in the proposal solution and the current best solution.
The extreme  case is a proposal where each 
node is in a separate connnected component.
Solving the fusion move with such a proposal
is equivalent to solving the original problem.
Solving the fusion move with a proposal with 
a single connected component will result 
in the current best solution.
Therefore the size of the proposals
should be small enough, such that solving
eq.~\ref{eq:edgeproblem} can
be done fast enough, but large enough to make
meaningful moves (??? this last sentence needs to be rewritten).

@@

\subsection{Properties and Polyhedral Interpretation}
An alternative interpretation of our method is shown in Fig.~\ref{fig:polyheadral}.
In each iteration the current and proposed segmentation define an inner polyheadral 
approximation of the original multicut polytope. 
Optimizing over this polytope is the same problem as the original one, but much smaller.
Furthermore, the costs does not change and and improvement in the smaller polytope will 
be the same in the original graph.
The choice of the proposal defines the shape of the auxiliary polytope. 
In the given toy example, the first (red) polytope give a huge improvement, the second proposal
defines the blue polytope which does not lead to an improvement. 
The third proposal generates the green polytope that includes the global optimal solution.

\begin{remark}[Optimality]
Let $y'$ be the current and $y''$ the proposed segmentation.
If the optimal solution $y^*$ lies in $P(G_{y' \OR y''})$ than a optimal fusion move will return $y^*$.
\end{remark}

>>>>

Fusion moves as defined in~\cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0^{\breve{y}}  & =  \{ ij \in E \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0^{\breve{y}}$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0^{\breve{y}} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{Lempitsky-2010}.
We iteratively improve the best solution by
fusing it with proposal solution.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.

@@

\begin{theorem}[Equivalence]\label{thm:Equivalence}
The back projection of the optimal segmentation $\bar{y}'$ of the contracted graph $G_y=(V_y,E_y, \bar{w})$
is an optimal solution of problem~\ref{eq:fusion_move_a}.
\begin{proof}
Let $y'$ be the back propagation of $\bar{y}'$, which is by definition a feasible for~\ref{eq:fusion_move_a}.
If $y'$ would not be an optimal solution than there must exists $y''$ with 
$\sum_{e\in E} w_ey_e'' > \sum_{e\in E} w_ey_e''$.
Since $y_e'$ and $y_e''$ are $0$ for all $e \in \mathcal{E}_0^{\breve{y}}$ we would have
\begin{align*}
  \sum_{\bar{e}\in E_y} \bar{w}_{\bar{e}}\bar{y}'_{\bar{e}}
 &= \sum_{e\in E \setminus \mathcal{E}_0^{\breve{y}}} w_ey_e'          
  = \sum_{e\in E} w_ey_e'' \\
 &> \sum_{e\in E} w_ey_e''
  = \sum_{e\in E \setminus \mathcal{E}_0^{\breve{y}}} w_ey_e''
  =\sum_{\bar{e}\in E_y} \bar{w}_{\bar{e}}\bar{y}''_{\bar{e}}
\end{align*}
where $\bar{y}''$ is the projection from $y''$ on $G_y$.
This contradicts that $y'$ is a optimal segmentation of $G_y$.
\end{proof}
%
\end{theorem}

@@

%-------------------------------------------------------------------------
\subsection{Polyhedral Interpretation}
A polyheadral interpretation of fusion moves is shown in Fig.~\ref{fig:polyheadral}.
In each iteration the current and proposed segmentation define an inner polyheadral 
approximation of the original polytope. This interpretation holds for original 
fusion moves~\cite{Lempitsky-2010} as well as for the proposed CCFusion.

In our case, optimizing over the inner polytope is the same kind of problem as the original multicut polytope, but much smaller.
Furthermore, the costs does not change and and improvement in the smaller polytope will 
be the same in the original graph, as shown in Theorem~\ref{thm:Equivalence}.

The choice of the proposal defines the shape of the auxiliary polytope. 
In the given toy example, the first (red) polytope give a huge improvement, the second proposal
defines the blue polytope which does not lead to an improvement. 
The third proposal generates the green polytope that includes the global optimal solution.

% \begin{remark}[Optimality]
% Let $y'$ be the current and $y''$ the proposed segmentation.
% If the optimal solution $y^*$ lies in $P(G_{y' \OR y''})$ than a optimal fusion move will return $y^*$.
% \end{remark}

@@

As discussed in \cite{Lempitsky-2010}, proposals
should have two properties: \emph{quality} 
and \emph{diversity}.
%
The quality of the individual proposals
ensures low energy solutions since the
result of the fusion cannot be higher then
the energy of the current best solution.
%
Diversity between the individual proposals
increase the chances to escape from local minima.
%
For correlation clustering fusion we add a third
property: \emph{size}.
The size of the contraction graph directly
depends on the number of connected components
in the proposal solution and the current best solution.
The extreme  case is a proposal where each 
node is in a separate connnected component.
Solving the fusion move with such a proposal
is equivalent to solving the original problem.
Solving the fusion move with a proposal with 
a single connected component will result 
in the current best solution.
Therefore the size of the proposals
should be small enough, such that solving
eq.~\ref{eq:edgeproblem} can
be done fast enough, but large enough to make
meaningful moves (??? this last sentence needs to be rewritten).

@@

%-----------------------------------------------------------------------
\subsection{Limitations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
